---
title: "Legacy Usage of `get_phrdw_data()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Legacy Usage of `get_phrdw_data()`}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE
)

# library(phrdwRdata)
# library(dplyr)
# library(forcats)
# library(tibble)
```

# Introduction

`phrdwRdata::get_phrdw_data()` function is designed to pull data for users in
an easy and efficient manner by abstracting away the nuances of database
connection, complexity of querying for OLAP (ie. data cubes) and relational
database (ie. SQL), and easy handling of some common filters, such as 
Health Authority Regions, etc.

The redesign of `phrdwRdata` package removes the restriction to work in
the Microsoft Client of R (ver 3.5.2), but respects *legacy* scripts by
retaining the order of arguments, and only updating the functions that are
deprecated in R version 4+. 

**This is an assurance to the users that they do not need to refactor their**
**scripts, and that their old scripts will continue to work.**

That said, it is *highly* recommended for users to update or upgrade their 
workflow to meet modern coding standards.

This vignette collects and revises (only if necessary) past `phrdwRdata`
tutorials and presents them all in one place.

# Prerequisites for legacy `phrdwRdata` package

The BCCDC Analytics Platform (Citrix) is running the required Microsoft R Client
(ver 3.5.2) that comes with `olapR` package and OLE DB connection files 
to drive OLAP (data cube) operations. No further action needed if using Citrix.

Users who are not using Citrix and want R installed locally will need to
ensure the following are met.

1.  Local PC is running 64-bit Windows.
2.  [Microsoft R Client 3.5.2][ms_client]. 

    - Alternatively, [CRAN R client 3.5.2][cran_client], but will require side-
    loading (ie. loading packages you did not install) `olapR` package which
    can be potentially problematic.

# CD Mart

Here's a table of all `CD Mart` datasets, keys used for joining, and a 
brief description if available.

## Non Tuberculosis Datasets

The `Investigation` dataset will retrieve non-identifiable client attributes
such as age and gender, along with Panorama data such as region, date, etiologic 
agent and stage. 

There are also support to retrieve linked datasets for `Risk Factor`s, Signs, 
`Symptom`s, and `UDF`s. At this time, only `Classification Group` == `Case` can
be pulled through the CD Mart datasets (what does this mean? is this true??? 
i don't see this in the code for CD).

**Note:** Only completed data is returned in the linked datasets. To see 
investigations where no Risk Factors, Signs & Symptoms or User Defined forms
where completed, a `LEFT JOIN` back to the `Investigation` dataset is required.

```{r}
#| label:   non-tb-datasets
#| eval:    false
#| echo:    false
#| message: false
#| include: false

phrdwRdata:::list_query_info$sql %>% 
  # dplyr::mutate(dataset_name = forcats::fct_inorder(dataset_name)) %>% 
  dplyr::mutate(dataset_name = as.factor(dataset_name)) %>%
  dplyr::filter(stringr::str_detect(dataset_name, '^TB', negate = T)) %>% 
  dplyr::group_by(dataset_name) %>% 
  dplyr::filter(
    stringr::str_detect(col, 'id$'),
    stringr::str_detect(.data$sql_func, 'join'),
  ) %>% 
  dplyr::summarise(col = list(unique(col))) %>% 
  dplyr::mutate(col = purrr::map_chr(col, paste, collapse = ', ')) %>% 
  dplyr::left_join(
    tibble::tribble(
      ~ dataset_name,         ~ Description,
      'Investigation',        paste("Retrieves Investigation and de-identified",
                                    "patient data."),
      'Client',               paste("Retrieves patients data with an option to",
                                    "include identifiers."),
      'Risk Factor',          paste("Retrieves Risk Factors that are marked",
                                    "Pertinent to an Investigation."),
      'Symptom',              paste("Retrieves Signs and Symptoms data."),
      'Observation',          paste("Retrieves the Observations from the Signs",
                                    "and Symptoms data."),
      'UDF',                  paste("Retrieves User Defined Forms data."),
      'Lab',                  paste("Retrieves Public Health System",
                                    "Laboratory data."),
      'Contacts',             paste("Retrieves Investigations based on",
                                    "the `Is A Contact Investigation` or",
                                    "`Ever A Contact Client` flags.",
                                    "Includes an option to retrieve",
                                    "patient and aboriginal identifiers."),
      'Transmission Events',  paste("Retrieves Transmission Events based",
                                    "on the source Investigation."),
      'Complication',         paste("Retrieves complication data"),
    )
  ) %>% 
  tidyr::replace_na(list(Description = '')) %>% 
  print
  dplyr::group_by(dataset_name) %>% 
  dplyr::mutate(
    dataset_name =
      dplyr::case_when(
        dplyr::row_number() >= 2 ~ '',
        TRUE ~ dataset_name
      )
  ) %>% 
  dplyr::rename(
    `Dataset Name` = dataset_name,
    `Linkable Key` = col
  ) %>% 
  knitr::kable()
  
```

|Dataset Name        |Linkable Key                                                                                |Description                                                                                                                                                             |
|:-------------------|:-------------------------------------------------------------------------------------------|:-----------------------------------------------------------------------------------------------------------------------------------------------------------------------|
|Client              |client_id                                                                                   |Retrieves patients data with an option to include identifiers.                                                                                                          |
|Complication        |investigation_id, encounter_id                                                              |Retrieves complication data                                                                                                                                             |
|Contacts            |client_id, disease_event_id, contact_disease_event_id                                       |Retrieves Investigations based on the `Is A Contact Investigation` or `Ever A Contact Client` flags. Includes an option to retrieve patient and aboriginal identifiers. |
|Investigation       |client_id, investigation_id, outbreak_id                                                    |Retrieves Investigation and de-identified patient data.                                                                                                                 |
|Lab                 |client_id, investigation_id, encounter_id, requisition_id, test_id, requisition_specimen_id |Retrieves Public Health System Laboratory data.                                                                                                                         |
|Observation         |disease_event_id                                                                            |Retrieves the Observations from the Signs and Symptoms data.                                                                                                            |
|Outbreaks           |investigation_id                                                                            |                                                                                                                                                                        |
|Risk Factor         |disease_event_id, risk_factor_id                                                            |Retrieves Risk Factors that are marked Pertinent to an Investigation.                                                                                                   |
|Symptom             |disease_event_id                                                                            |Retrieves Signs and Symptoms data.                                                                                                                                      |
|Transmission Events |disease_event_id                                                                            |Retrieves Transmission Events based on the source Investigation.                                                                                                        |
|UDF                 |disease_event_id, form_instance_disease_event_id                                            |Retrieves User Defined Forms data.                                                                                                                                      |

## Tuberculosis Datasets

```{r}
#| label:   tb-datasets
#| eval:    false
#| echo:    false
#| message: false

phrdwRdata:::list_query_info$sql %>% 
  # dplyr::mutate(dataset_name = forcats::fct_inorder(dataset_name)) %>% 
  dplyr::mutate(dataset_name = as.factor(dataset_name)) %>%
  dplyr::filter(stringr::str_detect(dataset_name, '^TB')) %>% 
  dplyr::group_by(dataset_name) %>% 
  dplyr::filter(
    stringr::str_detect(col, 'id$'),
    stringr::str_detect(.data$sql_func, 'join'),
  ) %>% 
  dplyr::summarise(col = list(unique(col))) %>% 
  dplyr::mutate(col = purrr::map_chr(col, paste, collapse = ', ')) %>% 
  dplyr::left_join(
    tibble::tribble(
      ~ dataset_name,           ~ Description,
      'TB Investigation',       paste('Retrieves TB specific investigation',
                                      'data. Includes an option to retrieve',
                                      'aboriginal fields if authorized.'),
      'TB Client',              paste('Retrieves TB specific patients',
                                      'data with an option to include',
                                      'identifiers.'),
      'TB Transmission Events', paste('Retrieves Transmission Events',
                                      'based on the source Investigation.'),
      'TB Contacts',            paste('Retrieves Investigations based',
                                      'on the `Is A Contact Investigation`',
                                      'or `Ever A Contact Client` flags.',
                                      'Includes an option to retrieve',
                                      'patient identifiers and aboriginal',
                                      'fields '),
      'TB TST Investigation',   paste('Retrieves TST data that is',
                                      'associated with an Investigation.'),
      'TB TST Client',          paste('Retrieves all TST data for a client,',
                                      'regardless of if the TST is',
                                      'associated with an Investigation.'),
      'TB Lab',                 paste('Retrieves TB specific laboratory data.'),
    )
  ) %>% 
  tidyr::replace_na(list(Description = '')) %>% 
  dplyr::group_by(dataset_name) %>% 
  dplyr::mutate(
    dataset_name =
      dplyr::case_when(
        dplyr::row_number() >= 2 ~ '',
        TRUE ~ dataset_name
      )
  ) %>% 
  dplyr::rename(
    `Dataset Name` = dataset_name,
    `Linkable Key` = col
  ) %>% 
  knitr::kable()
  
```

|Dataset Name           |Linkable Key                                                                                           |Description                                                                                                                                                                   |
|:----------------------|:------------------------------------------------------------------------------------------------------|:-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------|
|TB Client              |client_id                                                                                              |Retrieves TB specific patients data with an option to include identifiers.                                                                                                    |
|TB Contacts            |client_id, disease_event_id, contact_disease_event_id                                                  |Retrieves Investigations based on the `Is A Contact Investigation` or `Ever A Contact Client` flags. Includes an option to retrieve patient identifiers and aboriginal fields |
|TB Investigation       |client_id, disease_event_id                                                                            |Retrieves TB specific investigation data. Includes an option to retrieve aboriginal fields if authorized.                                                                     |
|TB Lab                 |client_id, investigation_id, encounter_id, requisition_id, test_id, result_id, requisition_specimen_id |Retrieves TB specific laboratory data.                                                                                                                                        |
|TB Transmission Events |disease_event_id                                                                                       |Retrieves Transmission Events based on the source Investigation.                                                                                                              |
|TB TST Client          |client_id, encounter_id                                                                                |Retrieves all TST data for a client, regardless of if the TST is associated with an Investigation.                                                                            |
|TB TST Investigation   |client_id, investigation_id, encounter_id                                                              |Retrieves TST data that is associated with an Investigation.                                                                                                                  |

## Script set up

1.  You will need to load or side-load the following:
```{r}
# Dependencies for `phrdwRdata`
library(dplyr)
library(digest)  
library(stringr)
library(RODBC)
library(RODBCext)

lib_loc <- "//Phsabc/root/BCCDC/Groups/Analytics_Resources/Coding/R/Library"
library(olapR,      lib.loc = log_loc) # if not using MS R Client
library(phrdwRdata, lib.loc = log_loc)
```

2.  Create a connection object
```{r}
phrdw_datamart <- 'CD Mart'
phrdw_datamart_connection <- phrdwRdata::connect_to_phrdw(phrdw_datamart)
```

3.  Set the start and end dates of your query
```{r}
query_start_date <- as.Date('2016-01-01')
query_end_date   <- as.Date('2018-01-01')
```

4.  Additional optional parameters

-  `disease`
-  `classification`
-  `surveillance_condition`
-  `surveillance_region_ha`

```{r}
surveillance_condition <- c("Legionella Infection")
classification         <- c("Confirmed",
                            "Confirmed Epi-Linked",
                            "Clinical",
                            "Probable")
surveillance_region_ha <- c("Interior",
                            "Fraser",
                            "Northern",
                            "Vancouver Coastal",
                            "Vancouver Island")
```

5.  Supply the above parameters in `get_phrdw_data()`

```{r}
dataset_name <- "Investigation"
investigation_dataset  <- 
  get_phrdw_data(
    phrdw_datamart_connection = phrdw_datamart_connection,
    phrdw_datamart            = phrdw_datamart, 
    dataset_name              = dataset_name, 
    query_start_date          = query_start_date, 
    query_end_date            = query_end_date,
    surveillance_condition    = surveillance_condition_vector,
    classification            = classification_vector,
    surveillance_region_ha    = surveillance_region_ha_vector
  )

dataset_name <- "UDF"
udf_dataset  <- 
  get_phrdw_data(
    phrdw_datamart_connection = phrdw_datamart_connection,
    phrdw_datamart            = phrdw_datamart, 
    dataset_name              = dataset_name, 
    query_start_date          = query_start_date, 
    query_end_date            = query_end_date,
    surveillance_condition    = surveillance_condition_vector,
    classification            = classification_vector,
    surveillance_region_ha    = surveillance_region_ha_vector
  )
```

## Arguments for identifiers and filters

```{r}
#| label:   args
#| eval:    false
#| echo:    false
#| message: false

require(tidyverse)

special_args_by_datasets <- 
  phrdwRdata:::list_query_info$sql %>% 
  filter(str_detect(check, 'id|user')) %>% 
  split(.$check)

special_args_by_datasets[c('indigenous_id', 'patient_id')] %>% 
  map(select, check, dataset_name) %>% 
  map_dfr(distinct) %>% 
  mutate(
    .keep = 'none',
    Argument =
      case_when(
        check == 'indigenous_id' ~ 'include_indigenous_identifiers',
        check == 'patient_id'    ~ 'include_patient_identifiers',
      ),
    Dataset = dataset_name
  ) %>% 
  group_by(Argument) %>% 
  mutate(
    Argument = 
      case_when(
        .default = Argument,
        row_number() > 1 ~ ''
      ),
    # Dataset = paste('-', Dataset)
  ) %>% 
  bind_rows(
  ) %>% 
  knitr::kable()
special_args_by_datasets$user %>% 
  drop_na(param_name) %>% 
  select(param_name, dataset_name) %>% 
  distinct %>% 
  arrange(param_name) %>% 
  group_by(param_name) %>% 
  summarise(
    dataset_name =
  )

  # summarise(dataset_name = list(dataset_name)) %>% 
  mutate(
    .keep = 'none',
    Argument = 
      case_when(
        .default = param_name,
        param_name == 'query_date' ~ 'query_(start|end)_date'
      ),
    Dataset = 
      map_chr(
        dataset_name, \(x) {
          n_check <- 
            length(unique(phrdwRdata:::list_query_info$sql$dataset_name))
          if (length(x) != n_check)
        }
      )
  ) 


```

|Argument                       |Dataset          |
|:------------------------------|:----------------|
|include_indigenous_identifiers |Contacts         |
|                               |TB Contacts      |
|                               |TB Investigation |
|include_patient_identifiers    |Investigation    |
|                               |Client           |
|                               |Contacts         |
|                               |TB Contacts      |
|                               |TB Client        |
|classification                 |All              |
|disease                        |All              |
|query_(start&#124;end)_date    |All              |
|surveillance_condition         |All              |
|surveillance_region_ha         |All              |

# Access to Data Cubes

## CDI Mart

[ms_client]: https://docs.microsoft.com/en-us/machine-learning-server/r-client/install-on-windows
[cran_client]: https://cran-archive.r-project.org/bin/windows/base/old/3.5.2
